The language of generalization: Probability, vagueness, and interaction

### 2018-03-24 paper/case2-habituals.Rmd
address comments in case 2

To test the generality of our theory, we use the same computational model and follow the same general experimental structure as in the first case study. 
We take the event analogue of prevalence to be the *rate* with which the event occurs (e.g., how often Mary smokes).^[
  Specifically, for the generalization "Mary smokes", the instances being generalized are *instances of Mary at different times*, and for an instance to have the feature, it means the instance of Mary is one in which she is smoking. \ndg{this footnote is confusing}
]
We test this by first measuring the prevalence (rate) prior distribution for various actions (e.g., how often different people smoke cigarettes; Expt. 2a).
We then measure endorsements of habitual statements while manipulating the referent prevalence  (Expt. 2b), and use our computational model to predict habitual endorsements. 
<!-- We estimate the priors using a novel data-analytic approach that takes advantage of the mixture-model validated in Expt. 1b. -->
By describing novel characters to participants, we are able to directly the manipulate the referent prevalence, which we were unable to do for generalizations about familiar categories.
<!-- In addition to directly manipulating the referent feature-probability, these experiments provide the first test of the generality of our theory: It can apply to generalizations about events as well as categories. -->

Finally, if habituals (and generics) are truly language for conveying generalizations, they should reflect speakers' expectations, not merely their observations. 
Our computational model predicts endorsement rates given a referent prevalence $p$. 
Does prevalence represent the actual, objective frequency in the world or the subjective, predictive belief in the head?
We elucidate this aspect of our theory by measuring participants' predictions about future rates and endorsements of habituals when causal forces intervene on the world (Expt. 2c). \ndg{give example of causal force intervening..}
We then use our model to test whether habitual endorsements are based on past frequencies or future expectations.
\ndg{possibly introduce this question briefly in intro, so readers are expecting it?}

@@

In order to generate model predictions for corresponding habitual statement, we first elicit the prior distributions over rates for different events.
For language about the behaviors of people, $P(p)$ represents a language user's background knowledge about the rates with which people do a behavior; this prior can be constructed as a distribution over *different people*, each of whom do the behavior with a different rate.
We designed our elicitation task to take advantage of the mixture-model representation of the prevalence prior used in Expt. 1.
In particular, we assume, to a first approximation, that the distribution over prevalence can be represented as a mixture of those who tend to do the action with a stable rate and those who do not do the action.^[
  It is likely that more than just these two possibilities are represented in people's intuitive prior, corresponding to individuals with additional traits or demographics.
  We assume here a simple two-component structure so as to not make the specification of the prior overly complex.
]
With the further assumption that, all else being equal, past is predictive of future behavior, we operationalize these two kinds of people as *people who have done the action before* and *people who have not the action before*.
We design this experiment to measure participants' beliefs about the relative proportion of these two kinds of people (as a measure of the mixture parameter in the model) as well as the rate at which people (who have done the action before) do the action (as a measure of the mean of stable-rate people \ndg{huh? stable rate people?}).
We will assume for simplicity that people who have never done the action before will rarely or never do the action.

@@

To construct our stimulus set, we choose actions from five categories of typical human behaviors: eating/consuming food and drug, work, wearing clothing, entertaining activities, and hobbies. 
For each category, we created pairs or triplets of events that shared a superordinate actions (e.g. *writing* poems vs. novels).
The events were chosen to intuitively cover a range of likely frequencies.
In total, thirty-one events were used.
For a full list the stimulus set, see Appendix D.

@@

\begin{enumerate}
\item ``How many \{men, women\} have \textsc{done action} before?'' \\

@@

\item ``For a typical \{man, woman\} who has \textsc{done action}  before, how frequently does he or she \textsc{do action}?''\\

@@

figure.habituals.priors.scatter <- 
  rs.habituals.bda.posterior.predictive.measurementSpace %>%
  select(-mean, -ci_lower, -ci_upper, -empirical_mean, -n, -cred_upper, -cred_lower) %>%
  spread(measurement, MAP) %>%
  ggplot(.,
       aes(x = `% people who have done action before`, 
           y = `Frequency of doing action (log scale)`, 
           color = gender))+
  scale_color_solarized()+
  geom_point()+
  #geom_errorbar(alpha = 0.3)+geom_errorbarh(alpha = 0.3)+
  geom_label_repel(data = . %>%
                    filter(item %in% example.habituals, gender == "female"), 
                  inherit.aes =F, segment.alpha = 1,label.size = 0.25,
                  segment.size = 5,
                  aes(x = `% people who have done action before`, 
           y = `Frequency of doing action (log scale)`, 
           label = item), force = 1, size = 2)+
  coord_fixed(1/7.5)+
  #xlab("Model posterior predictive")+
  #ylab("Empirical means")+
  theme(legend.position = c(0.83, 0.16))+
  scale_y_continuous(
      limits = c(-1.5, 6),
      breaks = c(0, 2.5, 4, 5.9),
      labels = c( "annually", "monthly", "weekly", "daily")
    )+
  scale_x_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))

@@

```{r habituals-prior-figure, fig.width = 15,fig.asp = 0.37, fig.cap="Prevalence priors for events (Expt. 2a). A: Mixture-model posterior predictions in comparison to the empirical mean responses for the two prior elicitation questions. Items cover the full spectrum of possible parameter values. B: Reconstructed prevalence priors. In order to show frequencies for events that are very rare across people (e.g., \"writes novels\"), extremely low frequencies (\\emph{almost never}) are omitted. Instead, height of the bars on left denote the Maximum A-Posteriori values of the mixture component (in terms of \\emph{proportion of people who have NEVER done action before} scatter-plot), reflecting the rarity of the event across people.", cache = F}
cowplot::plot_grid(
  figure.habituals.priors.scatter + theme(plot.margin = unit(c(6,6,6,0), "pt")), 
  figure.habituals.priors.marginals + theme(plot.margin = unit(c(6,0,6,6), "pt")),
  labels = c("A", "B"), #nrow = 1,
  align = 'h',
  rel_widths = c(1, 3)
  )
```

@@

The priors elicited cover a range of possible parameter values as intended (Figure \@ref(fig:priorScatter), scatter) \ndg{check fig ref}:
We observe a correlation in our items between the mean \% of Americans who have \textsc{done action} before (Question 1) and the mean log-frequency of action (Question 2) ($r_{1,2}(`r  n.items.habituals.prior.modeling`) = `r round(with(rs.habituals.bda.prior.summary.wide, cor(mix_mean, sfreq_mean)),2)`$).
Items in our data set that tend to be more popular actions also tend to be more frequent actions (e.g., \emph{wears socks}) and visa-versa (e.g., \emph{steals cars}), though there are notable exceptions (e.g., \emph{plays the banjo} is not popular but done frequently when done at all, as is \emph{smokes cigarettes}; \emph{goes to the movies} is a popular activity though not done particularly often).
This diversity is relevant because the speaker model (Eq. \ref{eq:S1}) will endorse habitual sentences (e.g., \emph{Sam goes to the movies vs. the ballet.}) contingent on these details of the prior distribution.

@@

Figure \@ref(fig:habituals-prior-figure) (right) shows example reconstructed priors.
In addition to specifying the correct way to combine our two prior-elicitation questions, using this inferred prior resolves two technical difficulties.
First, it smooths effects that are clearly results of the response format. 
For example, a very common rating for certain events is \emph{1 time per year}.
Presumably participants would be just as happy reporting \emph{approximately} 1 time per year (e.g., on average, 1.2 times per year); the raw data does not reflect this due to demands of the dependent measure.
Second, this methodology better captures the tails of the prior distribution (i.e., very frequent or very infrequent rates) which have relatively little data and need to be regularized by the analysis.

@@

mse.habituls.regression.freq.distinct <- compute_mse(d.hab.endorse.regression.freq.mix,
                                          "MAP_h", "prediction")

@@

Each event from Expt. 2a was then paired with between 2 - 4 frequencies, for which the habitual statement would be evaluated. 
Frequencies were always "3 times every \textsc{time period}" so that the referent character had always completed the event multiple times.
\textsc{time period}s were chosen specific to each event by simulating predictions of the endorsement model (Eq. \ref{eq:S1}) with the goal of maximizing the variability in endorsements across items.
\ndg{describe the motivation and procedure more clearly. it sounds a little fishy this way, but actually it's great: we set out to test the predictions of the model by choosing the items predicted to have most variability....}
For example, relatively high frequencies were chosen (e.g., three times every week vs. 2 months) for an item that was expected to occur rather often (e.g., "runs") because the model predicted that much of the variability in endorsement would occur in this range; for an item was expected to occur infrequently (e.g., "climbs mountains"), lower frequencies were chosen (e.g., three times every 5 years vs. 2 years).
In total, `r nrow(md.hab.long %>% select(habitual, time_period) %>% distinct())` unique items were created by pairing frequencies with events.

@@

On each trial, participants were presented with a \emph{past frequency statement} for a given event of the form: "In the past M \{weeks, months, years\}, \textsc{person} \textsc{did x} 3 times".
For example, "In the past month, Bill smoked cigarettes 3 times".
Participants were asked whether they agreed or disagreed with the corresponding habitual sentence: "\textsc{person does x}" (e.g.,"Bill smokes cigarettes.").
Participants completed thirty-seven trials, which were composed of the thirty-one items from the prior elicitation task randomly paired with either a male or female character name.
Six of these items were then also paired with a name of the opposite gender (e.g., participants rated both a female character and a male character who drank beer).
These were used for an exploratory analysis on differences in endorsements by gender of the target character.

@@

Habitual sentences were endorsed for a wide range of frequencies.
When actions are very infrequent (3 times in a 5-year interval), habituals can receive strong agreement (e.g., \emph{writes novels}, \emph{climbs mountains}).
When actions are relatively frequent (e.g., 3 times in a one month interval), habitual sentences can receive less than full endorsement (e.g., \emph{wears socks}, \emph{drinks coffee}). 
In our data, actions completed with a high frequency (3 times in a one week interval) receive at a minimum 75\% endorsement, though there is still variability among them (e.g., between 10-25\% of people disagree with \emph{wears a watch} and \emph{wears a bra}).
In addition, we observe that none of our items receive less than 25\% endorsement (i.e., a maximum of about 75\% of participants disagree with the felicity of the utterance), reflecting the fact that these statements are not altogether *false* even though the action is done very rarely.

@@

We now turn to our model-based analyses to better understand the endorsement data and the contribution of our model.
For all analyses, we collapse across gender of the referent character for endorsement judgments.
Parallel to our analysis of generic language endorsements, we articulate a set of simple regression models and a fixed-threshold alternative to our uncertain-threshold endorsement model.
Analyses which use the rate prior distribution $P(p_f)$ (all models except "referent frequency" regression) use a 50\% mixture of the inferred priors for each gender to construct a single prior distribution.
To model the uncertainty in the input measurements, we bootstrap the data for the regression models and we construct joint-inference, Bayesian data analytic models for the information-theoretic endorsement models (fixed-threshold and uncertain-threshold).

@@

To understand the role of frequency in habitual endorsement, we use the frequency supplied to the participants in our experiment as a predictor in a linear model. 
This model predicts the same endorsement level for two actions done with the same rate. 
Obvious counterexamples exist in our data set:
While participants are willing to endorse that a person "... climbs mountains" having done it 3 times in the past *year*, they are less willing to say that a person "... hikes" and not willing to say that a person "... runs".
Some actions done with a relative high rate (e.g., 3 times in a month) do not receive full endorsement (e.g., *smokes cigarettes*; Figure \@ref(fig:habituals-endorsement-figure) A).
Overall, the referent-frequency (in log-scale) predicts only a fraction of the variability in responses ($r^2(`r r2.hab.n`) = `r r2.habituls.regression.freq`$; MSE=$`r mse.habituls.regression.freq`$). 
In addition, for actions that are done on the time scale of years or longer (lower median of frequency), referent frequency no longer explains endorsements ($r^2(`r r2.hab.infrequent.n`) = `r r2.habituls.infrequent.regression.freq`$; MSE =$`r mse.habituls.infrequent.regression.freq`$).
The prevalence baseline does appreciably worse in this data set in comparison to the generics data set (Case Study 1) because we designed the stimulus set to be critical of a referent-frequency model. \ndg{i think it's not so much that we designed the stimuli to be critical, as that we were able to independently manipulate the referent-frequency from the prior, which we couldn't do for generics?}
<!-- An account purely based on the referent frequency does not explain endorsements in this data set. -->
<!-- Instead, we pursue an explanation where prior beliefs interact with an underspecified literal semantics to produce the pattern of habitual endorsement observed. -->

@@

In our empirically elicited priors, items differ the proportion of people who have done the action before (the *mixture parameter* of the mixture model; Figure \@ref(fig:habituals-prior-figure)).
This mixture parameter is a major contributor to the mean of the prevalence prior distribution, and thus is relates to the *cue validity* of a particular feature for a particular individual (see Appendix A for a derivation of the relationship between the prevalence prior and cue validity).
Thus, we take this parameter as an index of the *distinctiveness* of the action, analogous to *cue validity* in the case study of generic language.
We construct a regression model that treats endorsement as a linear combination of frequency and distinctiveness information. 
We use participants' responses to the question about the mixture parameter $\phi$ (i.e., the proportion of people who have done the action before) as well as frequency given to participants to predict habitual endorsement.

@@

This model is able to explain more of the variance in endorsements ($r^2(`r r2.hab.n`) = `r r2.habituls.regression.freq.distinct`$; MSE=$`r mse.habituls.regression.freq.distinct`$). 
It can differentiate events done with the same frequency (e.g., *writing poems* vs. *novels*, 3 times in the past 5 years), and increase endorsement of the more rare action (*novels*). 
Still, this model fails to capture fine-grained details in endorsement.
For example, going to the movies is a relatively nondistinctive action (many people do it) and going three times in a year is not very frequent, and yet people still strongly endorse the habitual $`r movies.year.data`$, while this regression model predicts quite lower judgments $`r movies.year.regression.distinct`$.
On the other hand, playing the banjo three times in the past two years is not strong evidence for the habitual, according to participants $`r banjo.2years.data`$.
Nevertheless, because playing the banjo is a distinct skill, the regression model wants to endorse the habitual strongly in this case $`r banjo.2years.regression.distinct`$.

@@

We next examine an information-theoretic endorsement model based on a fixed-threshold semantics.
A fixed-threshold model commits the habitual to conveying that a person *does the action with some rate*, and the threshold on rates is common to all actions.
As in Case Study 1, we incorporate this model into a Bayesian joint-inference model to infer the fixed-threshold and simultaneously predict both the priors data and the endorsement data (for more details on model implementation, see Appendix C).
We assume the referent-prevalence $p_{fk}$ being conveyed by the endorsement model (Eq. \ref{eq:S1}) is the frequency provided to participants (e.g., 3 times in the past year).
Additionally, to account for statements that would be literally false under this model (frequencies that fall below the fixed threshold), we include an additional noise parameter, as we did for the fixed-threshold model in Case Study 1. 
To learn about the credible values of the model's parameters and generate predictions given those inferred parameter values, we collected 2 MCMC chains of 100,000 iterations, discarding the first 50,000 iterations for burn in.

The data analytic model infers that a low threshold is likely: the inferred threshold in units of number of times per year is `r round(exp(fixed.threshold.hab.posterior[[1,"MAP"]]), 2)` [`r round(exp(fixed.threshold.hab.posterior[[1,"cred_lower"]]), 2)`,
`r round(exp(fixed.threshold.hab.posterior[[1,"cred_upper"]]), 2)`].
Compare this with the lowest referent-frequency used in our data set: $0.6$ times per year (3 times every 5 years).
Thus, all of the utterances evaluated under this fixed-threshold model were literally true.
As a result, the lowest endorsement this model can apply to an utterance is 0.5 (since both the habitual and silence are always true).
The fixed-threshold model exhibits a small dynamic range of endorsements, similar to the referent-frequency model.

A fixed-threshold utterance updates the interpreter model's beliefs differentially depending on the item.
For instance, it's highly unlikely that a random person *climbs mountains* with any non-zero rate; thus this model fully endorses "Mary climbs mountains" even with low-frequency (Figure \@ref(fig:habituals-endorsement-figure) A).
However, the model doesn't differentiate doing the same action with different frequencies.
It is equally true that a person does an action *with some frequency* for any frequency greater than zero, analogous to how "some Ks F" is equally true if 20\% or 50\% of Ks F.

Finally, the data-analytic model infers that `r 100*round(fixed.noise.hab.posterior[[1,"MAP"]], 2)`% [`r 100*round(fixed.noise.hab.posterior[[1,"cred_lower"]], 2)`,
`r 100*round(fixed.noise.hab.posterior[[1,"cred_upper"]], 2)`] of the data is noise and that the speaker optimality parameter is `r round(hab.fixed.threshold.s1opt[[1,"MAP"]], 2)` [`r round(hab.fixed.threshold.s1opt[[1,"cred_lower"]], 2)`,
`r round(hab.fixed.threshold.s1opt[[1,"cred_upper"]], 2)`].

\ndg{don't you want to report the posterior predictive correlation with data for this model?}

@@

We used the same data analytic approach for the uncertain threshold endorsement model and performed the same Bayesian statistical inference over the model to learn about its parameters and predictions. 
Again, this model has two fewer parameters than the fixed-threshold model (no fixed-threshold parameter and no extrinsic noise process).
As shown in Figure \@ref(fig:habituals-endorsement-figure)B, the uncertain-threshold endorsement model does a good job of accounting for the variability in responses ($r^2(`r r2.hab.n`) = `r r2.habituals.rsa.uncertain`$; MSE=$`r mse.habituals.rsa.uncertain`$), including actions done on the time scale of years or more  ($r^2(`r r2.hab.rsa.infrequent.n`) = `r r2.habituals.rsa.uncertain.infrequent`$; MSE=$`r mse.habituals.rsa.uncertain.infrequent`$).

>>>>

To test the generality of our theory, we use the same computational model and follow the same general experimental structure as in the first case study. 
We take the event analogue of prevalence to be the *rate* with which the event occurs (e.g., how often Mary smokes).^[
  Specifically, for the generalization "Mary smokes", the instances being generalized are *instances of Mary*.
]
We test this by first measuring the prevalence (rate) prior distribution for various actions (e.g., how often different people smoke cigarettes; Expt. 2a).
We then measure endorsements of habitual statements while manipulating the referent prevalence  (Expt. 2b), and use our computational model to predict habitual endorsements. 
<!-- We estimate the priors using a novel data-analytic approach that takes advantage of the mixture-model validated in Expt. 1b. -->
By describing novel characters to participants, we are able to directly the manipulate the referent prevalence, which we were unable to do for generalizations about familiar categories.
<!-- In addition to directly manipulating the referent feature-probability, these experiments provide the first test of the generality of our theory: It can apply to generalizations about events as well as categories. -->

Finally, if habituals (and generics) are truly language for conveying generalizations, they should reflect speakers' expectations, not merely their observations. 
Our computational model predicts endorsement rates given a referent prevalence $p$. 
Does prevalence represent the actual, objective frequency in the world (e.g., the rate at which a person has smoked cigarettes in the past) or a subjective, predictive belief in the head (e.g., the rate at which a person is expected to smoke in the future)?
Such a distinction could explain why "Supreme Court Justices have even social security numbers" sounds strange even if nine out of the nine current justices have even social security numbers [@Cohen1999]: Our predictions about the evenness of the *next* justice's social security number are driven by strong prior beliefs that selection for the Supreme Court is uncorrelated from the numerical properties of one's social security number; the current observations are not enough in this case to sway those beliefs.
<!-- not swayed by these observations due to -->
We elucidate this aspect of our theory by measuring participants' predictions about future rates and endorsements of habituals when causal forces intervene on the world (e.g., the person buys a pack of cigarettes; Expt. 2c).
We then use our model to test whether habitual endorsements are based on past frequencies or future expectations.

@@

In order to generate model predictions for corresponding habitual statement, we first elicit the prior distributions over rates for different events.
For language about the behaviors of people, $P(p)$ represents a language user's background knowledge about the rates with which people do a behavior; this prior can be constructed as a distribution over *different people*, each of whom do the behavior with a different rate.
We designed our elicitation task to take advantage of the mixture-model representation of the prevalence prior used in Expt. 1.
In particular, we assume, to a first approximation, that the distribution over prevalence can be represented as a mixture of those who tend to do the action with a stable rate and those who do not do the action.^[
  It is likely that more than just these two possibilities are represented in people's intuitive prior, corresponding to individuals with additional traits or demographics.
  We assume here a simple two-component structure so as to not make the specification of the prior overly complex.
]
With the further assumption that, all else being equal, past is predictive of future behavior, we operationalize these two kinds of people as *people who have done the action before* and *people who have not the action before*.
We design this experiment to measure participants' beliefs about the relative proportion of these two kinds of people (as a measure of the mixture parameter in the model) as well as the rate at which people (who have done the action before) do the action.
We will assume for simplicity that people who have never done the action before will probably never do the action.

@@

To construct our stimulus set, we choose actions from five categories of typical human behaviors: eating/consuming food and drug, work, wearing clothing, entertaining activities, and hobbies. 
For each category, we created pairs or triplets of events that shared a superordinate actions (e.g. *writing* poems vs. novels).
The events were chosen to intuitively cover a range of likely frequencies.
In total, thirty-one events were used.
For a full list of the stimuli used in Expts. 2a-c, see Appendix D.

@@

\begin{enumerate}
\item ``How many \{men, women\} have *done action* before?'' \\

@@

\item ``For a typical \{man, woman\} who has *done action*  before, how frequently does he or she \textsc{do action}?''\\

@@

rs.habituals.bda.posterior.predictive.measurementSpace.wide <- rs.habituals.bda.posterior.predictive.measurementSpace %>%
  select(-mean, -ci_lower, -ci_upper, -empirical_mean, -n, -cred_upper, -cred_lower) %>%
  spread(measurement, MAP)

figure.habituals.priors.scatter <- ggplot(
  rs.habituals.bda.posterior.predictive.measurementSpace.wide %>%
                    filter(item %in% example.habituals, gender == "female"),
       aes(x = `% people who have done action before`, 
           y = `Frequency of doing action (log scale)`, 
           color = gender, label = item))+
  scale_color_solarized()+
  geom_point(data = rs.habituals.bda.posterior.predictive.measurementSpace.wide %>%
                    filter(!(item %in% example.habituals) | gender == "male"),
             inherit.aes = F, 
             aes(
               x = `% people who have done action before`, 
               y = `Frequency of doing action (log scale)`, color = gender))+
  geom_point()+
  geom_label_repel(segment.alpha = 1,label.size = 0.25,
                   force = 10, size = 2, color = 'black')+

  #geom_errorbar(alpha = 0.3)+geom_errorbarh(alpha = 0.3)+
  coord_fixed(1/7.5)+
  #xlab("Model posterior predictive")+
  #ylab("Empirical means")+
  theme(legend.position = c(0.83, 0.16))+
  scale_y_continuous(
      limits = c(-1.5, 6),
      breaks = c(0, 2.5, 4, 5.9),
      labels = c( "annually", "monthly", "weekly", "daily")
    )+
  scale_x_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))

@@

```{r habituals-prior-figure, fig.width = 15,fig.asp = 0.37, fig.cap="Prevalence priors for events (Expt. 2a). A: Parameters of prevalence priors for the 31 items in Expt. 2. Items cover much of the range of possible parameter values. B: Reconstructed prevalence priors. In order to show frequencies for events that are very rare across people (e.g., \"writes novels\"), extremely low frequencies (\\emph{almost never}) are omitted. Instead, height of the bars on left denote the Maximum A-Posteriori values of the mixture component (in terms of \\emph{proportion of people who have never done action before} scatter-plot), reflecting the (inverse) popularity of the event across people.", cache = F}
cowplot::plot_grid(
  figure.habituals.priors.scatter + theme(plot.margin = unit(c(6,6,6,0), "pt")), 
  figure.habituals.priors.marginals + theme(plot.margin = unit(c(6,0,6,6), "pt")),
  labels = c("A", "B"), #nrow = 1,
  align = 'h',
  rel_widths = c(1, 3)
  )
```

@@

The priors elicited cover a range of possible parameter values as intended (Figure\ \@ref(fig:habituals-prior-figure)A):
We observe a correlation in our items between the mean \% of Americans who have \textsc{done action} before (Question 1) and the mean log-frequency of action (Question 2) ($r_{1,2}(`r  n.items.habituals.prior.modeling`) = `r round(with(rs.habituals.bda.prior.summary.wide, cor(mix_mean, sfreq_mean)),2)`$).
Items in our data set that tend to be more popular actions also tend to be more frequent actions (e.g., \emph{wears socks}) and visa-versa (e.g., \emph{steals cars}), though there are notable exceptions (e.g., \emph{plays the banjo} is not popular but done frequently when done at all, as is \emph{smokes cigarettes}; \emph{goes to the movies} is a popular activity though not done particularly often).
This diversity is relevant because the speaker model (Eq. \ref{eq:S1}) will endorse habitual sentences (e.g., \emph{Sam goes to the movies vs. the ballet.}) contingent on these details of the prior distribution.

@@

Figure\ \@ref(fig:habituals-prior-figure)B shows example reconstructed priors.
In addition to specifying the correct way to combine our two prior-elicitation questions, using this inferred prior resolves two technical difficulties.
First, it smooths effects that are clearly results of the response format. 
For example, a very common rating for certain events is \emph{1 time per year}.
Presumably participants would be just as happy reporting \emph{approximately} 1 time per year (e.g., on average, 1.2 times per year); the raw data does not reflect this due to demands of the dependent measure.
Second, this methodology better captures the tails of the prior distribution (i.e., very frequent or very infrequent rates) which have relatively little data and need to be regularized by the analysis.

@@

mse.habituls.regression.freq.distinct <- compute_mse(d.hab.endorse.regression.freq.mix,
                                          "MAP_h", "prediction")
```

@@

Each event from Expt. 2a was then paired with between two to four frequencies, for which the habitual statement would be evaluated. 
Frequencies were presented in terms of a character performing the action "three times in the past *time interval*".
We chose to always have the character perform the action three times to provide a strong test of a baseline hypothesis that the habitual encodes the person has done the action multiple times in the past.

Different time intervals were chosen for each event in order to maximize the variability of responses within each item. 
Specifically, we used the endorsement model to generate predictions based on the prior elicitation data (Expt. 2a) for each item, and chose between two and four time intervals across which maximal variability was predicted. 
For example, relatively high frequencies were chosen (e.g., time intervals of weeks and months) for items expected to occur rather often (e.g., *runs*); for an item was expected to occur infrequently (e.g., *climbs mountains*), lower frequencies were chosen (e.g., time intervals of years or longer) because the model predicted that much of the variability in endorsement would occur in that range.
In total, `r nrow(m.hab.endorse.rsa %>% distinct(habitual, time_period))` unique items were created by pairing frequencies with events. 
The full list of items and frequencies can be found in Appendix D.

@@

On each trial, participants were presented with a *past frequency statement* for a given event of the form: "In the past \{week, 2 weeks, month, 5 years, ...\}, *Person did X* 3 times".
For example, "In the past month, Bill smoked cigarettes 3 times".
Participants were asked whether they agreed or disagreed with the corresponding habitual sentence: "Person does X" (e.g.,"Bill smokes cigarettes.").
Participants completed thirty-seven trials, which were composed of the thirty-one items from the prior elicitation task randomly paired with either a male or female character name.
Six of these items were then also paired with a name of the opposite gender (e.g., participants rated both a female character and a male character who drank beer).
These were used for an exploratory analysis on differences in endorsements by gender of the target character.

@@

Habitual sentences were endorsed for a wide range of frequencies.
When actions are very infrequent (3 times in a 5-year interval), habituals can receive strong agreement (e.g., \emph{writes novels}, \emph{climbs mountains}).
When actions are relatively frequent (e.g., 3 times in a one month interval), habitual sentences can receive less than full endorsement (e.g., \emph{wears socks}, \emph{drinks coffee}). 
In our data, actions completed with a high frequency (3 times in a one week interval) receive at a minimum 75\% endorsement, though there is still variability among them (e.g., between 10-25\% of people disagree that people who wore a watch or wore a bra 3 times in the past week \emph{wears a watch} or \emph{wears a bra} habitually).
In addition, we observe that none of our items receive less than 25\% endorsement (i.e., a maximum of about 75\% of participants disagree with the habitual utterances), reflecting the fact that these statements are not altogether *false* even though the action may be done very rarely.

@@

We now turn to our model-based analyses to better understand the endorsement data and the contribution of our model.
For all analyses, we collapse across gender of the referent character for endorsement judgments.
Parallel to our analysis of generic language endorsements, we articulate a set of simple regression models and a fixed-threshold alternative to our uncertain-threshold endorsement model.
Analyses which use the prevalence prior distribution $P(p)$ (all models except "referent frequency" regression) use a 50\% mixture of the inferred priors for each gender to construct a single prevalence prior distribution.
Parallel to our analysis of generics model, we model uncertainty in the input measurements by bootstrapping the data for the regression models and constructing joint-inference, Bayesian data analytic models for the information-theoretic endorsement models (fixed-threshold and uncertain-threshold).

@@

To understand the role of frequency in habitual endorsement, we use the frequency supplied to the participants in our experiment as a predictor in a linear model. 
This model predicts the same endorsement level for two actions done with the same rate. 
Obvious counterexamples exist in our data set:
While participants are willing to endorse that a person "... climbs mountains" having done it 3 times in the past *year*, they are less willing to say that a person "... hikes" and not willing to say that a person "... runs".
Some actions done with a relative high rate (e.g., 3 times in a month) do not receive full endorsement (e.g., *smokes cigarettes*; Figure \@ref(fig:habituals-endorsement-figure)A).
Overall, the referent-frequency (in log-scale) predicts only a fraction of the variability in responses ($r^2(`r r2.hab.n`) = `r r2.habituls.regression.freq`$; MSE=$`r mse.habituls.regression.freq`$). 
In addition, for actions that are done on the time scale of years or longer (lower median of frequency), referent frequency no longer explains endorsements ($r^2(`r r2.hab.infrequent.n`) = `r r2.habituls.infrequent.regression.freq`$; MSE =$`r mse.habituls.infrequent.regression.freq`$).
The prevalence baseline does appreciably worse in this data set in comparison to the generics data set (Case Study 1) because we were able to independently manipulate the referent-frequency separate from the prevalence priors, which we could not do for generics about familiar categories.

<!-- designed the stimulus set to be critical of a referent-frequency model. \ndg{i think it's not so much that we designed the stimuli to be critical, as that we were able to independently manipulate the referent-frequency from the prior, which we couldn't do for generics?} -->
<!-- An account purely based on the referent frequency does not explain endorsements in this data set. -->
<!-- Instead, we pursue an explanation where prior beliefs interact with an underspecified literal semantics to produce the pattern of habitual endorsement observed. -->

@@

In our empirically elicited priors, items differ the proportion of people who have done the action before (the *mixture parameter* of the mixture model; Figure \@ref(fig:habituals-prior-figure)A x-axis).
This mixture parameter is a major contributor to the mean of the prevalence prior distribution, and thus is relates to the *cue validity* of a particular feature for a particular individual (see Appendix A for a derivation of the relationship between the prevalence prior and cue validity).
Thus, we take this parameter as an index of the *distinctiveness* of the action, analogous to *cue validity* in the case study of generic language.
We construct a regression model that treats endorsement as a linear combination of frequency and distinctiveness information. 
We use participants' responses to the question about the mixture parameter $\phi$ (i.e., the proportion of people who have done the action before) as well as frequency given to participants to predict habitual endorsement.

@@

This model is able to explain more of the variance in endorsements ($r^2(`r r2.hab.n`) = `r r2.habituls.regression.freq.distinct`$; MSE=$`r mse.habituls.regression.freq.distinct`$). 
It can differentiate events done with the same frequency (e.g., *writing poems* vs. *novels*, 3 times in the past 5 years) by increasing endorsement of the more rare action (*novels*). 
Still, this model fails to capture fine-grained differences in endorsement.
For example, going to the movies is a relatively nondistinctive action (many people do it) and going three times in a year is not very frequent, and yet people still strongly endorse the habitual ($`r movies.year.data`$), while this regression model predicts quite lower judgments ($`r movies.year.regression.distinct`$).
On the other hand, playing the banjo three times in the past two years is not strong evidence for the habitual, according to participants ($`r banjo.2years.data`$).
Nevertheless, because playing the banjo is a distinct skill, the regression model wants to endorse the habitual strongly in this case ($`r banjo.2years.regression.distinct`$).

@@

We next examine an information-theoretic endorsement model based on a fixed-threshold semantics.
A fixed-threshold model commits the habitual to conveying literally that a person *does the action with some non-zero frequency*, and the threshold on frequency is the same for all actions.
As in Case Study 1, we incorporate this model into a Bayesian joint-inference model to infer the fixed-threshold and simultaneously predict both the priors data and the endorsement data (for more details on model implementation, see Appendix C).
We assume the referent-prevalence $p$ being conveyed by the endorsement model (Eq. \ref{eq:S1}) is the frequency provided to participants (e.g., 3 times in the past year).
Additionally, to account for statements that would be literally false under this model (frequencies that fall below the fixed threshold), we include an additional noise parameter, as we did for the fixed-threshold model in Case Study 1. 
To learn about the credible values of the model's parameters and generate predictions given those inferred parameter values, we collected 2 MCMC chains of 100,000 iterations, discarding the first 50,000 iterations for burn in.

The data analytic model infers that a low threshold is likely: the Maximum A-Posteriori threshold and 95\% credible interval in units of number of times per year is `r round(exp(fixed.threshold.hab.posterior[[1,"MAP"]]), 2)` [`r round(exp(fixed.threshold.hab.posterior[[1,"cred_lower"]]), 2)`,
`r round(exp(fixed.threshold.hab.posterior[[1,"cred_upper"]]), 2)`].
Compare this with the lowest referent-frequency used in our data set: $0.6$ times per year (3 times every 5 years).
Thus, all of the utterances evaluated under this fixed-threshold model were literally true.
As a result, the lowest endorsement this model can apply to an utterance is 0.5 (since both the habitual and silence are always true).
The fixed-threshold model exhibits a small dynamic range of endorsements, similar to the referent-frequency model (Figure\ \@ref(fig:habituals-endorsement-figure)B).

The fixed-threshold habitual updates the interpreter model's prior beliefs differentially depending on the item.
For instance, because of the distinctiveness of *climbs mountains*, the fixed-threshold endorsement model fully endorses the habitual ("Mary climbs mountains") even at a low frequency (Figure \@ref(fig:habituals-endorsement-figure)A).
However, the model cannot differentiate among different frequencies of doing the same action.
It is equally true that a person does an action with non-zero frequency for any frequency greater than zero, analogous to how "Some dogs are friendly" is equally true whether 20\% or 50\% or 80\% of dogs are friendly.

Overall, the fixed-threshold model is able to predict only a fraction of the variance in human endorsements ($r^2(`r r2.hab.n`) = `r r2.habituals.rsa.fixed`$; MSE=$`r mse.habituals.rsa.fixed`$).
The model does this by inferring that `r 100*round(fixed.noise.hab.posterior[[1,"MAP"]], 2)`% [`r 100*round(fixed.noise.hab.posterior[[1,"cred_lower"]], 2)`,
`r 100*round(fixed.noise.hab.posterior[[1,"cred_upper"]], 2)`] of the data is noise and that the speaker optimality parameter is `r round(hab.fixed.threshold.s1opt[[1,"MAP"]], 2)` [`r round(hab.fixed.threshold.s1opt[[1,"cred_lower"]], 2)`,
`r round(hab.fixed.threshold.s1opt[[1,"cred_upper"]], 2)`].

@@

We used the same data analytic approach for the uncertain threshold endorsement model and performed the same Bayesian statistical inference over the model to learn about its parameters and predictions. 
Again, this model has two fewer parameters than the fixed-threshold model (no fixed-threshold parameter and no extrinsic noise process).
As shown in Figure\ \@ref(fig:habituals-endorsement-figure)B, the uncertain-threshold endorsement model does a good job of accounting for the variability in responses ($r^2(`r r2.hab.n`) = `r r2.habituals.rsa.uncertain`$; MSE=$`r mse.habituals.rsa.uncertain`$), including actions done on the time scale of years or more  ($r^2(`r r2.hab.rsa.infrequent.n`) = `r r2.habituals.rsa.uncertain.infrequent`$; MSE=$`r mse.habituals.rsa.uncertain.infrequent`$).
